### 2015-05-28 papers/histogram/thesis/thesis.tex

\subsection{The Wang-Landau (WL) method}
\label{sec:wang_landau}

@@

    \item accept the move $s_i\to s_f$ with probability
      \begin{align*}
        P_m\p{s_i\to s_f} =\max\set{\f{\p{\tilde T_d\p{E_i,\Delta
                E}+c}/\p{n_i+c}} {\p{\tilde T_d\p{E_f,-\Delta
                E}+c}/\p{n_f+c}}, \exp\p{-\f{\Delta E}{kT_{\t{min}}}}}
      \end{align*}
      for a predetermined number $c$ and temperature $T_{\t{min}}$.
      \label{alg:tmmc_prob}
    \end{enumerate}

@@

Aside from a minimum temperature $T_{\t{min}}$ and an end condition,
Algorithm \ref{alg:tmmc} contains one free parameter, $c$, used in
step \ref{alg:tmmc_prob} to modify (\ref{eq:tmmc_prob_simplified}).
This parameter makes the rejection of transitions more conservative
when there are too few statistics in the histogram $\tilde T_d$ to be
confident in the current estimate of the transition matrix. In order
to make the effects of $c$ negligible as $\tilde T_d$ collects more
statistics throughout initialization, $c$ should be of order unity,
though its optimal value is generally system-dependent. In our
implementation of Algorithm \ref{alg:tmmc}, we use $c=16$.

@@

While the Wang-Landau method comes equipped with an end condition of
its own, most histogram methods require some external means of
deciding when they are finished initializing. Wang-Landau's
``natural'' end condition does not advantage it over other methods,
however, as the WL end condition makes use of a free parameter (the
cutoff $c$), which any end condition will require. In order to compare
all methods fairly, we need a method-independent end condition to
initialization. The most straightforward of such end conditions is to
simply to stop initializing after some fixed number of Monte Carlo
iterations, and finish up any other calculations in the algorithm if
appropriate. We thus initialize systems of $N$ spheres for $10^5N^2$
iterations; the choice of $10^5N^2$ we address in Section
\ref{sec:results}. In general, however, one might want an end
condition which is independent of system size. We came up with a four
such end conditions as examples, though we do not use them in
simulations analyzed in this paper.

@@

\begin{figure}[p]
  \centering
  \includegraphics[width=0.75\textwidth]{figs/internal-energy.pdf}
  \caption[Internal energy]
  {Specific internal energy curves averaged over 30 $N=25$
    simulations.}
  \label{fig:internal_energy}
\end{figure}

@@

\begin{figure}[p]
  \centering
  \includegraphics[width=0.7\textwidth]{figs/u-scaling.pdf}
  \caption[Internal energy error scaling]
  {Specific internal energy error scaling.}
  \label{fig:u_scaling}
\end{figure}

\begin{figure}[p]
  \centering
  \includegraphics[width=0.75\textwidth]{figs/cv-scaling.pdf}
  \caption[Heat capacity error scaling]
  {Specific heat capacity error scaling.}
  \label{fig:cv_scaling}
\end{figure}

\begin{figure}[p]
  \centering
  \includegraphics[width=0.7\textwidth]{figs/s-scaling.pdf}
  \caption[Configurational entropy error scaling]
  {Specific configurational entropy error scaling.}
  \label{fig:s_scaling}
\end{figure}

\begin{figure}[p]
  \centering
  \includegraphics[width=0.75\textwidth]{figs/u-comps.pdf}
  \caption[Internal energy error comparisons]
  {Specific internal energy errors, plotted against those from
    converged flat histogram weight simulations.}
  \label{fig:u_comps}
\end{figure}

\begin{figure}[p]
  \centering
  \includegraphics[width=0.7\textwidth]{figs/cv-comps.pdf}
  \caption[Heat capacity error comparisons]
  {Specific heat capacity errors, plotted against those from converged
    flat histogram weight simulations.}
  \label{fig:cv_comps}
\end{figure}

\begin{figure}[p]
  \centering
  \includegraphics[width=0.7\textwidth]{figs/s-comps.pdf}
  \caption[Configurational entropy error comparisons]
  {Specific configurational entropy errors, plotted against those from
    converged flat histogram weight simulations.}
  \label{fig:s_comps}
\end{figure}

Figures \ref{fig:u_comps}--\ref{fig:s_comps} provide a different
visualization of the same information as in Figures
\ref{fig:u_scaling}--\ref{fig:s_scaling}, plotting the errors of all
histogram methods against those of converged flat histogram weight
simulations. The fact that some points fall below the diagonal
reference line is strange, and indicates that some regular simulations
are outperforming the converged flat ones. We do not understand why
such events occur, as converged flat simulations should be using
high-quality golden data to construct weight arrays, and should
therefore outperform all standard simulations.

>>>>

\subsection{The Wang-Landau method}
\label{sec:wang_landau}

@@

    \item accept the move $s_i\to s_f$ with probability
      \begin{align*}
        P_m\p{s_i\to s_f}=\max\set{\f{\p{\tilde T_d\p{E_i,\Delta E}+c}
            /\p{n_i+c}} {\p{\tilde T_d\p{E_f,-\Delta E}+c}/\p{n_f+c}},
          \exp\p{-\f{\Delta E}{kT_{\t{min}}}}}
      \end{align*}
      for a predetermined number $c$ and temperature $T_{\t{min}}$.
      \label{alg:tmmc_prob}
    \end{enumerate}

@@

Aside from a minimum temperature $T_{\t{min}}$ and an end condition,
Algorithm \ref{alg:tmmc} contains one free parameter, $c$, used in
step \ref{alg:tmmc_prob} to modify (\ref{eq:tmmc_prob_simplified}).
This parameter, which is our own addition to the algorithm provided in
\cite{tmmc}, makes the rejection of transitions more conservative when
there are too few statistics in the histogram $\tilde T_d$ to be
confident in the current estimate of the transition matrix. Without
this parameter, the TMMC initialization routine can get stuck at
energies which, due to poor collection of statistics in $\tilde T_d$,
appear to be more important than they actually are. In order to make
the effects of $c$ negligible as $\tilde T_d$ collects more statistics
throughout initialization, $c$ should be of order unity, though its
optimal value is generally system-dependent. In our implementation of
Algorithm \ref{alg:tmmc}, we use $c=16$.

@@

While the Wang-Landau method comes equipped with an end condition of
its own, most histogram methods require some external means of
deciding when they are finished initializing. Wang-Landau's
``natural'' end condition does not advantage it over other methods,
however, as the Wang-Landau end condition makes use of a free
parameter (the cutoff $c$), which any end condition will require. In
order to compare all methods fairly, we need a method-independent end
condition to initialization. The most straightforward of such end
conditions is to simply to stop initializing after some fixed number
of Monte Carlo iterations, and finish up any other calculations in the
algorithm if appropriate. We thus initialize systems of $N$ spheres
for $10^5N^2$ iterations; the choice of $10^5N^2$ we address in
Section \ref{sec:results}. In general, however, one might want an end
condition which is independent of system size. We came up with a four
such end conditions as examples, though we do not use them in
simulations analyzed in this paper.

@@

\begin{figure}[!b]
  \centering
  \includegraphics[width=0.75\textwidth]{figs/internal-energy.pdf}
  \caption[Internal energy]
  {Specific internal energy curves averaged over 30 $N=25$
    simulations.}
  \label{fig:internal_energy}
\end{figure}

@@

\begin{figure}[!p]
  \centering
  \includegraphics[width=0.7\textwidth]{figs/u-scaling.pdf}
  \caption[Internal energy error scaling]
  {Specific internal energy error scaling.}
  \label{fig:u_scaling}
\end{figure}

\begin{figure}[!p]
  \centering
  \includegraphics[width=0.75\textwidth]{figs/cv-scaling.pdf}
  \caption[Heat capacity error scaling]
  {Specific heat capacity error scaling.}
  \label{fig:cv_scaling}
\end{figure}

\begin{figure}[!p]
  \centering
  \includegraphics[width=0.7\textwidth]{figs/s-scaling.pdf}
  \caption[Configurational entropy error scaling]
  {Specific configurational entropy error scaling.}
  \label{fig:s_scaling}
\end{figure}

\begin{figure}[!p]
  \centering
  \includegraphics[width=0.75\textwidth]{figs/u-comps.pdf}
  \caption[Internal energy error comparisons]
  {Specific internal energy errors, plotted against those from
    converged flat histogram weight simulations.}
  \label{fig:u_comps}
\end{figure}

\begin{figure}[!p]
  \centering
  \includegraphics[width=0.7\textwidth]{figs/cv-comps.pdf}
  \caption[Heat capacity error comparisons]
  {Specific heat capacity errors, plotted against those from converged
    flat histogram weight simulations.}
  \label{fig:cv_comps}
\end{figure}

\begin{figure}[!p]
  \centering
  \includegraphics[width=0.7\textwidth]{figs/s-comps.pdf}
  \caption[Configurational entropy error comparisons]
  {Specific configurational entropy errors, plotted against those from
    converged flat histogram weight simulations.}
  \label{fig:s_comps}
\end{figure}

Figures \ref{fig:u_comps}--\ref{fig:s_comps} provide a different
visualization of the same information as in Figures
\ref{fig:u_scaling}--\ref{fig:s_scaling}, plotting the errors of all
histogram methods against those of converged flat histogram weight
simulations. The fact that some points fall below the diagonal
reference line is strange, and indicates that some regular simulations
are outperforming the converged flat ones. We do not understand why
such events occur, as converged flat simulations should be using high
quality golden data to construct weight arrays, and should therefore
outperform all standard simulations.
